---
title: "Arrival distribution and Simulatuin data"
author: "Group 39"
date: "6/1/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Libraries
we will need the following libraries: 
```{r libraries, include=TRUE,echo=TRUE}

library(readxl)
library(chron)
library(fitdistrplus)
library(magrittr)
library(lubridate)

```

# 1. Fit the data to our needs

## 1.1  Empirical Data
Empirical data - observations. 
Here we read the date from an excel file that provided to us.
In the file there are the arrival time stamps of the first 50 voters in a Big Kalpi and Accessibility Kalpi.
```{r generate, include=TRUE,echo=TRUE}

sf <- read.csv('voting schedule.csv', header = TRUE)
sf1 <- sf[,1]
sf2 <- sf[,2]

```


## 1.2  Define function 
We don't want the arrival time stamp but the gap between two arrivals.
```{r Define, include=TRUE,echo=FALSE}

arrivalDiff <- function(arrival) {
  diff <- c()
  for (i in 2:length(arrival)) {
    diff <-
      append(diff,
             times(arrival[i], format = "h:m:s") - times(arrival[i - 1], format ="h:m:s"))
  }
  return(diff)
}

```


## 1.3  Assing data
Now we need to format the data into usable data in a integer, hour format.
```{r Assing, include=TRUE,echo=FALSE}

bigKalpi <- arrivalDiff(sf1)
accessibleKalpi <- arrivalDiff(sf2)

x <- strptime(bigKalpi, format = "%H:%M:%S")
numericBigKalpi <- (hour(x)*3600 + minute(x) * 60 + second(x))/60

x <- strptime(accessibleKalpi, format = "%H:%M:%S")
numericAccessibleKalpi <- (hour(x)*3600 + minute(x) * 60 + second(x))/60

```


# 2. Fitting

## 2.1 Getting a sence of how the data distributes like 
Let's see how Big Kalpi looks:
```{r Getting a sence, include=TRUE,echo=TRUE}

sample <- numericBigKalpi
par(mfrow = c(1, 3)) # defines 2 plots on the same window
plot(sample, ylab = "Sample values")
hist(sample, xlab="seconds", ylab = "Frequency ")
boxplot(sample)

```
from the first look, we think that the data did not came from a normal distribution-from the histogram, maby from exponential distribution.
look's like there are no outliers. and just in case we did not figure them out, we want to keep them because the data might came from a real data collection.
And how the Accessible Kalpi look like:
```{r , include=TRUE,echo=TRUE}

sample <- numericAccessibleKalpi
par(mfrow = c(1, 3)) # defines 2 plots on the same window
plot(sample, ylab = "Sample values")
hist(sample, xlab="seconds", ylab = "Frequency ")
boxplot(sample)

```
Here it's more clear that the data is more likely got from an exponential distribution. boxcox cannot be trusted, because it belong to normal distribution.
we can notice that there are some outliers, but we descided to keep them.

## 2.2 Trying to fit Big Kalpi 
```{r fit Big Kalpi, include=TRUE,echo=TRUE}

sample <- numericBigKalpi

ExpFit <- fitdist(sample, "exp") 
summary(ExpFit) 

GammaFit <- fitdist(sample, "gamma")
summary(GammaFit)

```

### 2.2.1 Goodness-of-fit statistics for Big Kalpi 
Now let's look even further, we will plot the data into a histogram (theoretical density), CDF, QQ-plot and PP-plot.
```{r}

par(mfrow = c(2, 2))
legendText <- c("exp", "gamma") # define legend text
denscomp(list(ExpFit, GammaFit), legendtext = legendText) # histograms
qqcomp(list(ExpFit, GammaFit), legendtext = legendText)  # qq plot
cdfcomp(list(ExpFit, GammaFit), legendtext = legendText) # cdf plot
ppcomp(list(ExpFit, GammaFit), legendtext = legendText) # pp plot

ExpGOF <- gofstat(ExpFit) %>% print()
ExpGOF$kstest %>% print()

```
the data more likely to be from a exponential distribution.
## 2.3 Trying to fit Accessibility Kalpi 
```{r fit Accessibility Kalpi , include=TRUE,echo=TRUE}

sample <- numericAccessibleKalpi

ExpFit <- fitdist(sample, "exp") 
summary(ExpFit) 

GammaFit <- fitdist(sample, "gamma")
summary(GammaFit)

```

### 2.3.2 Goodness-of-fit statistics for Accessibility Kalpi 
```{r}

par(mfrow = c(2, 2))
legendText <- c("exp", "gamma") # define legend text
denscomp(list(ExpFit, GammaFit), legendtext = legendText) # histograms
qqcomp(list(ExpFit, GammaFit), legendtext = legendText)  # qq plot
cdfcomp(list(ExpFit, GammaFit), legendtext = legendText) # cdf plot
ppcomp(list(ExpFit, GammaFit), legendtext = legendText) # pp plot

ExpGOF <- gofstat(ExpFit) %>% print()
ExpGOF$kstest %>% print()

```
the data more likely to be from a exponential distribution.

# CONCLUTION
both distributions were not rejected in both Big and Accessibility Kalpis, we can assume the following:

Big Kalpi is distributed exponentially with lambda = 1.775148	

Accessible Kalpi is also distributed exponentially with lambda = 0.445699


\newpage


# Analising the KPIs for the current situation

## 1. Import the data
This is the data regarding the first alternative
```{r, include=TRUE,echo=TRUE}

data0 <- read.csv("data0.csv",header = T)[,2:3]
voters0 <- data0$vote_to_left_prob0*100
renege0 <- data0$reneged_from_left_prob0*100

```

## 2. Fitting

### 2.1.1 Fitting Voter to a Normal distribution
We must fit the data to a normal distribution. Let's see if it works.
```{r, include=TRUE,echo=TRUE}

NormFitV <- fitdist(voters0, "norm") 
summary(NormFitV) 

```

### 2.1.2 Goodness-of-fit statistics for Voter
Now let's look even further, we will plot the data into a histogram (theoretical density), CDF, QQ-plot and PP-plot.
```{r}

par(mfrow = c(2, 2))
denscomp(NormFitV, legendtext = "norm") # histograms
qqcomp(NormFitV, legendtext = "norm")  # qq plot
cdfcomp(NormFitV, legendtext = "norm") # cdf plot
ppcomp(NormFitV, legendtext = "norm") # pp plot

```
there are two outliers, and we will remove them. so, we have to remove the same rows in order to be able to use CRN in further analysis. 

```{r}
xx <- voters0[voters0!=(voters0)[11] & voters0!=(voters0)[12]]
print(shapiro.test(xx))
```
according to Shapiro-Wilk normality test:
the p-value is 0.2 and above 0.05. So we now know our variable voters0 is normally distributed.

### 2.2.1 Fitting Renege to a Normal distribution
```{r, include=TRUE,echo=TRUE}

renege0 <- renege0[renege0!=(renege0)[11] & renege0!=(renege0)[12]]

NormFitR <- fitdist(renege0, "norm")
summary(NormFitR)

```

### 2.2.2 Goodness-of-fit statistics for Renege
```{r}

par(mfrow = c(2, 2))
denscomp(NormFitR, legendtext = "norm") # histograms
qqcomp(NormFitR, legendtext = "norm")  # qq plot
cdfcomp(NormFitR, legendtext = "norm") # cdf plot
ppcomp(NormFitR, legendtext = "norm") # pp plot

print(shapiro.test(renege0))

```

according to Shapiro-Wilk normality test:
the p-value is 0.53 and above 0.05. So we now know our variable renege0 is normally distributed.

## CONCLUTION
Both data frames hypothesis has not been rejected, therefor they are both distributed Normally. 


\newpage


# Analising the KPIs for the first alternative

## 1. Import the data
This is the data regarding the first alternative
```{r, include=TRUE,echo=TRUE}

data1 <- read.csv("data1.csv",header = T)[,2:3]
voters1 <- data1$vote_to_left_prob1*100
renege1 <- data1$reneged_from_left_prob1*100

renege1 <- renege1[renege1!=(renege1)[11] & renege1!=(renege1)[12]]
voters1 <- voters1[voters1!=(voters1)[11] & voters1!=(voters1)[12]]

```


## 2. Fitting

### 2.1.1 Fitting Voter to a Normal distribution
We must fit the data to a normal distribution. Let's see if it works.
```{r, include=TRUE,echo=TRUE}

NormFitV <- fitdist(voters1, "norm") 
summary(NormFitV) 

```

### 2.1.2 Goodness-of-fit statistics for Voter
Now let's look even further, we will plot the data into a histogram (theoretical density), CDF, QQ-plot and PP-plot.
```{r}

par(mfrow = c(2, 2))
denscomp(NormFitV, legendtext = "norm") # histograms
qqcomp(NormFitV, legendtext = "norm")  # qq plot
cdfcomp(NormFitV, legendtext = "norm") # cdf plot
ppcomp(NormFitV, legendtext = "norm") # pp plot

print(shapiro.test(voters1))

```
there is two points that looks like outliers, but we will keep them.
according to Shapiro-Wilk normality test:
The data is normal if the p-value is above 0.05. So we now know our variable voters1 is normally distributed.

### 2.2.1 Fitting Renege to a Normal distribution
```{r, include=TRUE,echo=TRUE}

NormFitR <- fitdist(renege1, "norm")
summary(NormFitR)

```

### 2.2.2 Goodness-of-fit statistics for Renege
```{r}

par(mfrow = c(2, 2))
denscomp(NormFitR, legendtext = "norm") # histograms
qqcomp(NormFitR, legendtext = "norm")  # qq plot
cdfcomp(NormFitR, legendtext = "norm") # cdf plot
ppcomp(NormFitR, legendtext = "norm") # pp plot

print(shapiro.test(renege1))

```
we thinks here that there are not outliers.
according to Shapiro-Wilk normality test:
the p-value is 0.76 and above 0.05. So we now know our variable renege1 is normally distributed.

## CONCLUTION
Both data frames hypothesis has not been rejected, therefor they are both distributed Normally. 


\newpage


# Analising the KPIs for the second alternative

## 1. Import the data
This is the data regarding the first alternative
```{r, include=TRUE,echo=TRUE}

data2 <- read.csv("data2.csv",header = T)[,2:3]
voters2 <- data2$vote_to_left_prob2*100
renege2 <- data2$reneged_from_left_prob2*100

renege2 <- renege2[renege2!=(renege2)[11] & renege2!=(renege2)[12]]
voters2 <- voters2[voters2!=(voters2)[11] & voters2!=(voters2)[12]]

```

## 2. Fitting

### 2.1.1 Fitting Voter to a Normal distribution
We must fit the data to a normal distribution. Let's see if it works.
```{r, include=TRUE,echo=TRUE}

NormFitV <- fitdist(voters2, "norm") 
summary(NormFitV) 

```

### 2.1.2 Goodness-of-fit statistics for Voter
Now let's look even further, we will plot the data into a histogram (theoretical density), CDF, QQ-plot and PP-plot.
```{r}

par(mfrow = c(2, 2))
denscomp(NormFitV, legendtext = "norm") # histograms
qqcomp(NormFitV, legendtext = "norm")  # qq plot
cdfcomp(NormFitV, legendtext = "norm") # cdf plot
ppcomp(NormFitV, legendtext = "norm") # pp plot

print(shapiro.test(voters2))

```
no outliers here.
according to Shapiro-Wilk normality test:
The data is normal if the p-value is above 0.05. So we now know our variable voters2 is normally distributed.

### 2.2.1 Fitting Renege to a Normal distribution
```{r, include=TRUE,echo=TRUE}

NormFitR <- fitdist(renege2, "norm")
summary(NormFitR)

```

### 2.2.2 Goodness-of-fit statistics for Renege
```{r}

par(mfrow = c(2, 2))
denscomp(NormFitR, legendtext = "norm") # histograms
qqcomp(NormFitR, legendtext = "norm")  # qq plot
cdfcomp(NormFitR, legendtext = "norm") # cdf plot
ppcomp(NormFitR, legendtext = "norm") # pp plot

print(shapiro.test(renege2))

```

according to Shapiro-Wilk normality test:
the p-value is 0.07 and above 0.05. So we now know our variable renege2 is normally distributed.


\newpage


# Analising the KPIs for the third alternative

## 1. Import the data
This is the data regarding the first alternative
```{r, include=TRUE,echo=TRUE}

data3 <- read.csv("data3.csv",header = T)[,2:3]
voters3 <- data3$vote_to_left_prob3*100
renege3 <- data3$reneged_from_left_prob3*100

renege3 <- renege3[renege3!=(renege3)[11] & renege3!=(renege3)[12]]
voters3 <- voters3[voters3!=(voters3)[11] & voters3!=(voters3)[12]]

```

## 2. Fitting

### 2.1.1 Fitting Voter to a Normal distribution
We must fit the data to a normal distribution. Let's see if it works.
```{r, include=TRUE,echo=TRUE}

NormFitV <- fitdist(voters3, "norm") 
summary(NormFitV) 

```

### 2.1.2 Goodness-of-fit statistics for Voter
Now let's look even further, we will plot the data into a histogram (theoretical density), CDF, QQ-plot and PP-plot.
```{r}

par(mfrow = c(2, 2))
denscomp(NormFitV, legendtext = "norm") # histograms
qqcomp(NormFitV, legendtext = "norm")  # qq plot
cdfcomp(NormFitV, legendtext = "norm") # cdf plot
ppcomp(NormFitV, legendtext = "norm") # pp plot

print(shapiro.test(voters3))

```
generally, no outliers here.
according to Shapiro-Wilk normality test:
The data is normal if the p-value is above 0.05. So we now know our variable voters3 is normally distributed.

### 2.2.1 Fitting Renege to a Normal distribution
```{r, include=TRUE,echo=TRUE}

NormFitR <- fitdist(renege3, "norm")
summary(NormFitR)

```

### 2.2.2 Goodness-of-fit statistics for Renege
```{r}

par(mfrow = c(2, 2))
denscomp(NormFitR, legendtext = "norm") # histograms
qqcomp(NormFitR, legendtext = "norm")  # qq plot
cdfcomp(NormFitR, legendtext = "norm") # cdf plot
ppcomp(NormFitR, legendtext = "norm") # pp plot

print(shapiro.test(renege3))

```
we can detect two points that's a little far from the line in qqplot, but we will keep them.
according to Shapiro-Wilk normality test:
the p-value is 0.49 and above 0.05. So we now know our variable vote_to_left_prob1 is normally distributed.

## CONCLUTION
Both data frames hypothesis has not been rejected, therefor they are both distributed Normally. 


## Now we can get the final data:

```{r}
data0 <- read.csv("data0.csv",header = T)[,2:3]
data0 <- data0[-c(11,12),]
write.csv(data0, "data0ready.csv")

data1 <- read.csv("data1.csv",header = T)[,2:3]
data1 <- data1[-c(11,12),]
write.csv(data1, "data1ready.csv")

data2 <- read.csv("data2.csv",header = T)[,2:3]
data2 <- data2[-c(11,12),]
write.csv(data2, "data2ready.csv")

data3 <- read.csv("data3.csv",header = T)[,2:3]
data3 <- data3[-c(11,12),]
write.csv(data3, "data3ready.csv")

```
